Flowchart of Current Algorithm

### 2009-05-18 docs/flowchart/flowchart.tex
minor fixes to description

$\Ub$     & total velocity ($\Ub = \Ubt + w_0 \er$) & cm~s$^{-1}$ \\

@@

$\etarho$ & $\etarho \equiv \overline{(\rho' \Ubt \cdot \er)}$ & g~cm$^{-2}$~s$^{-1}$ \\

@@

\item For any quantity, $\phi$, we define $\overline{\phi} = ${\bf Avg}$(\phi)$
as the average over a layer at constant radius r, $\Omega_H$, as
\begin{equation}
\overline{\phi(r)}
= \frac{1}{\mathrm{A}(\Omega_H)}\int_{\Omega_H} \phi(r,{\bf x}) \; dA \enskip .
\end{equation}
\begin{description}
\item[planar:] We simply average cells at constant height.
\item[spherical:] See the description in paper IV, section 2.2 
(or section \ref{Sec:Avg} in these notes).
\end{description}

\item {\bf React State}$[\rho^{\inp},(\rho h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp}, 
p_0^{\inp}]\rightarrow [\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, 
(\rho \omegadot_k)^{\outp}, (\rho\Hnuc)^{\outp}]$
is the process by which we evolve the species and enthalpy from 
$X_k^{\inp} \rightarrow X_k^{\outp}$ 
and $(\rho h)^{\inp} \rightarrow (\rho h)^{\outp}$ by solving the following 
system of equations over a time interval of  $\Delta t/2$,
\begin{eqnarray}
\frac{\partial X_k}{\partial t} &=& \omegadot_k\enskip,\\
\frac{\partial (\rho h)}{\partial t} &=& \rho \Hnuc + \rho \Hext\enskip .
\end{eqnarray}
  In particular, to evolve the species, we solve the system:
\begin{eqnarray}
\frac{dX_k}{dt} &=& \omegadot_k(\rho,X_k,T)\enskip, \\
\frac{dT}{dt} &=&\frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right )
\enskip. \label{eq:reacttemp}
\end{eqnarray}
using the stiff ordinary differential equation integration methods
provided by the {\tt VODE} package (cite vode).  The absolute error
tolerances are set to $10^{-12}$ for the species, and a relative
tolerance of $10^{-5}$ is used for the temperature.  The integration
yields the new values of the mass fractions, $X_k^{\outp}$.  Equation
(\ref{eq:reacttemp}) is derived from the full temperature evolution equation
from equation (8) in paper III,
\begin{equation}
\label{eq:tempenthalpy}
\frac{DT}{Dt} = \frac{1}{\rho c_p} \left[ \left(1 - \rho h_p\right) \frac{D p}{D t}
 - \sum_k \rho \xi_k {\omegadot}_k 
 + \rho \Hnuc + \rho \Hext \right] \enskip , 
\end{equation}
by assuming that the pressure is constant during the burn state.  In
evolving these equations, we need to evaluate $c_p$ and $\xi_k$.  In
theory, this means evaluating the equation of state for each
right-hand side evaluation that {\tt VODE} requires.  In practice, we
freeze $c_p$ and $\xi_k$ at the start of the integration time step and
compute them using $\rho^{\inp}, X_k^{\inp},$ and $T^{\inp}$ as inputs
to the equation of state.  We update temperature using
$T^{\outp} = T(\rho^\outp,h^\outp,X_k^\outp)$ if {\tt use\_tfromp = FALSE} or 
$T^{\outp} = T(\rho^\outp,p_0,X_k^\outp)$  if {\tt use\_tfromp = TRUE}.  
Note that the density remains unchanged
during the {\bf React State} step, i.e., $\rho^{\outp} = \rho^{\inp}$.

@@

\item[planar:] We discretize the base state density evolution 
equation from paper III equation (29) without the $\etarho$ term,
\begin{equation}
\frac{\partial \rhozero}{\partial t} = - \nablab \cdotb \left( \rhozero w_0 \er \right)
\enskip ,
\label{eq:rho0upd_new}
\end{equation}
to compute the new base state density,
\begin{equation}
\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp} - \frac{\dt}{\dr} 
\left [ \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j+\myhalf} -  
         \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j-\myhalf} \right ] \enskip,
\end{equation}
where $j$ refers to the one-dimensional index in the radial direction.
The interface states, $\rho_0^{\outp,\nph}$, are computed using the
procedure described in paper II, Appendix A.\\

@@

We discretize the base state enthalpy equation,
\begin{equation}
\frac{\partial (\rho h)_0}{\partial t} = -\nabla\cdot\left[(\rho h)_0w_0\er\right] 
+ \psi \enskip ,
\end{equation}
to compute the new base state enthalpy,
\begin{equation}
(\rho h)_{0,j}^{\outp} = (\rho h)_{0,j}^{\inp} - \frac{\dt}{\Delta r}
\left\{ \left[ (\rho h)_0^{\nph} w_0^{\inp}\right]_{j+\myhalf} -  
         \left[ (\rho h)_0^{\nph} w_0^{\inp}\right]_{j-\myhalf} \right\}
+ \dt\psi_j^{\inp} \enskip ,
\end{equation}
where the interface states $(\rho h)_0^{n+\myhalf}$, are again found via the procedure 
described in paper II, Appendix A.\\

@@

The local velocity field is described by paper III equation (37),
\begin{equation}
\frac{\partial\Ubt}{\partial t} = 
- \left(\Ubt+w_0\right) \cdotb \nablab \Ubt
- \left(\Ubt \cdotb \er\right) \frac{\partial w_0}{\partial r} \er
- \frac{1}{\rho} \nablab\pi
+ \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} \er
- \frac{(\rho-\rhozero)}{\rho} \; g \; \er  \label{eq:utildeupd}  \enskip .
\end{equation}

We construct time-centered velocities at edges, $\uadvone$, 
using a second-order unsplit Godunov procedure followed by an elliptic
solve, as described in Appendix B of paper III.  We note that 
$\uadvone$ satisfies the discrete versions of paper III equation (21),
\begin{equation}
\int_{\Omega_H} \Ubt \cdotb \er \; dA = 0 \label{eq:udoterzero} \enskip ,
\end{equation}
and paper III equation (39),
\begin{equation}
\nablab \cdotb (\beta_0 \Ubt )  = \beta_0 \left(S - \Sbar \right )\enskip ,
\label{eq:tildeconstraint}
\end{equation}
with the discretization
\begin{equation}
\nablab \cdotb \left(\beta_0^n \uadvone\right) = 
\beta_0^n \left(S^{\nph,\star} - \overline{S^{\nph,\star}}\right) \enskip .
\end{equation}
The discretization with the volume discrepancy correction is
\begin{equation}
\nablab \cdotb \left(\beta_0^n \uadvone\right) = 
\beta_0^n \left\{ \left(S^{\nph,\star} - \overline{S^{\nph,\star}}\right)
+ \frac{f}{\gammabar^n p_0^n}
\left[\frac{p(\rho,h,X_k)^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n}\right]\right\} \enskip .
\end{equation}

@@

%--------------------------------------------------------------------------
% STEP 4
%--------------------------------------------------------------------------
\item[Step 4.] {\em Advect the base state, then the full state, through a time interval 
of $\dt.$}

@@

\item \label{step:species update} 
Update the species.  Here we consider only the advection terms, 
neglecting the reaction terms, and solve a discretized version of
\begin{equation}
\frac{\partial (\rho X_k)}{\partial t} + \nablab \cdotb (\Ub \rho X_k) = 0 \enskip .
\label{eq:species}
\end{equation}
The update consists of two steps:

@@

  \item Compute the species edge states, $(\rho X_k)^{(1),\nph,\star}$,
  for the conservative update of $(\rho X_k)^{(1)}$. 
  Here we predict $\rho^{'(1)} = \rho^{(1)} - \rhozero^n$ and 
  $X_k^{(1)} = (\rho X_k)^{(1)} / \rho^{(1)}$ to time-centered edges to obtain 
  $\rho^{'(1),\nph,\star}$ and $X_k^{(1),\nph,\star}$ using a second-order
  unsplit Godunov procedure, as described in paper II,
  Appendix A, using $\Vb = \uadvone+w_0^{\nph,\star} \er$.  Convert the 
  perturbational edge state to a full state using
\begin{equation}
\rho^{(1),\nph,\star} = 
\rho^{'(1),\nph,\star} + \frac{\rho_0^n + \rho_0^{(2a),\star}}{2} \enskip .
\label{eq:rhoXedgestate}
\end{equation}
  In planar geometry, we use 4th-order spatial averaging to compute $\rho_0$ on 
  radial faces.  In spherical geometry, we first map $\rho_0$ to Cartesian cell-centers 
  using the mapping in section \ref{Sec:1D Cell-Centered to Cartesian Cell-Centered}, 
  and then use 2nd-order spatial averaging to compute $\rho_0$ on all faces.

  \item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2),\star}$
  without explicitly including the reaction terms,
\begin{equation}
(\rho X_k)^{(2),\star} = (\rho X_k)^{(1)} 
 - \dt \; \left\{ \nablab \cdotb \left[ \left(\uadvone+w_0^{\nph,\star} \er\right)  
  (\rho X_k)^{(1),\nph,\star} \right] \right\} \enskip ,
\end{equation}
\begin{equation}
\rho^{(2),\star} = \sum_k (\rho X_k)^{(2),\star} \enskip ,
\end{equation}
\begin{equation}
X_k^{(2),\star} = (\rho X_k)^{(2),\star} / \rho^{(2),\star}
\end{equation}

@@

\begin{description}
\item For planar geometry,
\begin{equation}
 \etarho^{\nph,\star} =  {\rm {\bf Avg}} \sum_k \left[ \left(\uadvone \cdotb \er + w_0^{\nph,\star}\right) (\rho X_k)^{(1),\nph,\star} \right] - w_0^{\nph,\star} \rho_0^{\nph,\star,{\rm pred}} \enskip ,
\end{equation}
\item For spherical geometry, first construct 
$\eta_{\rho}^{{\rm cart},\nph,\star} = [\rho'(\Ubt\cdot\eb_r)]^{n+\myhalf,\star}$ using:
\begin{equation}
\eta_{\rho}^{{\rm cart},\nph,\star} = \left[\left(\frac{\rho^{(1)}+\rho^{(2),\star}}{2}\right)-\left(\frac{\rho_0^n+\rho_0^{(2a),\star}}{2}\right)\right] \sum_d \left(\uadvone \cdot \eb_d\right)n_d.
\end{equation}
Then, $\etarho^{\nph,\star}$ is the cell-centered average of $\eta_{\rho}^{{\rm cart},\nph,\star}$,
\begin{equation}
\etarho^{\nph,\star} = \overline{\eta_{\rho}^{{\rm cart},\nph,\star}}.
\end{equation}
This gives a cell-centered $\etarho^{\nph,\star}$.  To get $\etarho^{\nph,\star}$ at 
radial edges, average the two neighboring radial cell-centers.
\end{description}

@@

\item Update the enthalpy.  Here we consider only the advective terms, 
neglecting the reaction terms and thermal diffusion terms, and solve a discretized version of
  \begin{equation}
  \frac{\partial (\rho h)}{\partial t}  = - \nablab \cdotb (\Ub \rho h)
+ \psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r} 
\enskip , \label{eq:rhohupdadv} 
  \end{equation}
For spherical geometry, experience has shown that solving the analytically equivalent form,
\begin{equation}
  \frac{\partial (\rho h)}{\partial t}  = - \nablab \cdotb (\Ub \rho h)
+ \psi + \nabla \cdot (\Ubt p_0) - p_0 \nabla \cdot \Ubt  
\enskip , \label{eq:rhohupdadv2} 
\end{equation}
results in a better solution.

@@

  \item Compute the enthalpy edge state, $(\rho h)^{(1),\nph,\star},$
    for the conservative update of $(\rho h)^{(1)}.$  Here we predict 
    $(\rho h)' = (\rho h)^{(1)} - (\rho h)_0$ to time-centered edges to obtain 
    $(\rho h)^{'(1),\nph,\star}$, 
    using a second-order unsplit Godunov procedure, 
    as described in paper II, Appendix A, using $\Vb =
    \uadvone+w_0^{\nph,\star} \er$.  We do not include the reaction
    terms in the enthalpy prediction, since we accounted for them
    already in {\bf React State}.  Convert the perturbational edge
    state to a full state using

@@

  \begin{eqnarray}
  (\rho h)^{(2),\star} &=& (\rho h)^{(1)} - \dt \; \left\{ \nablab
      \cdotb \left[ \left(\uadvone+w_0^{\nph,\star} \er\right) (\rho
      h)^{(1),\nph,\star} \right] \right\} \nonumber \\ && + \dt \;
    \left(\uadvone \cdotb \er\right) \left(\frac{\partial
      p_0}{\partial r} \right)^{n} + \dt \; \psi^{\nph,\star} \enskip ,
  \end{eqnarray}

@@

  \begin{eqnarray}
  (\rho h)^{(2),\star} &=& (\rho h)^{(1)} - \dt \; \left\{ \nablab
      \cdotb \left[ \left(\uadvone+w_0^{\nph,\star} \er\right) (\rho
      h)^{(1),\nph,\star} \right] \right\} \nonumber \\ 
    && + \dt \; \left \{ \nabla \cdot \left (\uadvone p_0^{n} \right ) 
       - p_0^{n} \nabla \cdot \uadvone \right \} \nonumber \\
    && + \dt \; \psi^{\nph,\star} \enskip ,
  \end{eqnarray}

@@

The procedure to construct $\uadvtwo$ is described in detail in Appendix B of paper III
and is analogous to the procedure used in step 2, but with updated values
for $w_0$ and $\pizero.$  We note that $\uadvtwo$ satisfies the discrete versions of 
equations (\ref{eq:udoterzero}) and (\ref{eq:tildeconstraint}), with the discretization
\begin{equation}
\nablab \cdotb \left(\beta_0^{\nph} \uadvtwo\right) =
\beta_0^{\nph}\left(S^{\nph} - \overline{S^{\nph}}\right) \enskip ,
\end{equation}
\begin{equation}
\beta_0^{\nph} = \frac{ \beta_0^n +  \beta_0^{n+1,\star} }{2}; ~~~~~
 \beta_0^{n+1,\star} = \beta \left(\rho_0^{n+1,\star}, p_0^{n+1,\star}, \gammabar^{n+1,\star}, g^{n+1,\star}\right) \enskip ,
\end{equation}
where $\beta_0$ is computed as described in paper III, Appendix C.

The discretization with the volume discrepancy correction is
\begin{equation}
\nablab \cdotb \left(\beta_0^{\nph} \uadvtwo\right) =
\beta_0^{\nph}\left\{\left(S^{\nph} - \overline{S^{\nph}}\right)
 + \frac{f}{\gammabar^{\nph} p_0^{\nph}}\cdot 
\left[\frac{p(\rho,h,X_k)^{\nph} - \overline{p(\rho,h,X_k)^{\nph}}}{\Delta t^n}\right]\right\} \enskip ,
\end{equation}

%--------------------------------------------------------------------------
% STEP 8
%--------------------------------------------------------------------------
\item[Step 8.] {\em Advect the base state, then the full state, through a time interval of $\dt.$}

@@

\item Compute the species edge states, $(\rho X_k)^{(1),n+\myhalf}$, 
using $\Vb = \uadvtwo+w_0^{\nph}\er$.  Convert the perturbational edge 
state to a full state using 
\begin{equation}
\rho^{(1),\nph} = \rho^{'(1),\nph} + \frac{\rho_0^n + \rho_0^{(2a)}}{2}\enskip .
\end{equation}

\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2)}$ using
\begin{equation}
(\rho X_k)^{(2)} = (\rho X_k)^{(1)} 
- \dt \; \left\{ \nablab \cdotb \left[\left(\uadvtwo+w_0^{\nph} \er\right)  
(\rho X_k)^{(1),\nph} \right] \right\}  \enskip ,
\end{equation}
\begin{equation}
\rho^{(2)} = \sum_k (\rho X_k)^{(2)} \enskip ,
\end{equation}
\begin{equation}
X_k^{(2)} = (\rho X_k)^{(2)} / \rho^{(2)}
\end{equation}

@@

\item For planar geometry,
\begin{equation}
 \etarho^{\nph} = {\rm {\bf Avg}} \sum_k \left [\left(\uadvtwo \cdotb \er + w_0^{\nph}\right) (\rho X_k)^{(1),\nph} \right] - w_0^{\nph} \rho_0^{\nph,{\rm pred}} \enskip ,
\end{equation}
\item For spherical geometry, first construct 
$\eta_{\rho}^{{\rm cart},\nph} = [\rho'(\Ubt\cdot\eb_r)]^{n+\myhalf}$ using:
\begin{equation}
\eta_{\rho}^{{\rm cart},\nph} = \left[\left(\frac{\rho^{(1)}+\rho^{(2)}}{2}\right)-\left(\frac{\rho_0^n+\rho_0^{(2a)}}{2}\right)\right] \sum_d \left(\uadvtwo \cdot \eb_d\right)n_d.
\end{equation}
Then, $\etarho^{\nph}$ is the cell-centered average of $\eta_{\rho}^{{\rm cart},\nph}$,
\begin{equation}
\etarho^{\nph} = \overline{\eta_{\rho}^{{\rm cart},\nph}}.
\end{equation}
This gives a cell-centered $\etarho^{\nph}$.  To get $\etarho^{\nph}$ at 
radial edges, average the two neighboring cell-centers.
\end{description}

>>>>

$\Ub$     & total velocity ($\Ub = \Ubt + w_0 \eb_r$) & cm~s$^{-1}$ \\

@@

$\etarho$ & $\etarho \equiv \overline{(\rho' \Ubt \cdot \eb_r)}$ & g~cm$^{-2}$~s$^{-1}$ \\

@@

\item For any quantity, $\phi$, we define $\overline{\phi} = ${\bf Avg}$(\phi)$
as the average over a layer at constant radius r, $\Omega_H$, as
\begin{equation}
\overline{\phi(r)}
= \frac{1}{\mathrm{A}(\Omega_H)}\int_{\Omega_H} \phi(r,{\bf x}) \; dA \enskip .
\end{equation}
\begin{description}
\item[planar:] This is a straightforward average since the radial coordinate is 
in alignment with the lateral coordinate.
\item[spherical:] See the description in paper IV, section 2.2 
(or section \ref{Sec:Avg} in these notes).
\end{description}

\item {\bf React State}$[\rho^{\inp},(\rho h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp}, 
p_0^{\inp}]\rightarrow [\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, 
(\rho \omegadot_k)^{\outp}, (\rho\Hnuc)^{\outp}]$
is the process by which we evolve the species and enthalpy from 
$X_k^{\inp} \rightarrow X_k^{\outp}$ 
and $(\rho h)^{\inp} \rightarrow (\rho h)^{\outp}$ by solving the following 
system of equations over a time interval of  $\Delta t/2$,
\begin{eqnarray}
\frac{\partial X_k}{\partial t} &=& \omegadot_k\enskip,\\
\frac{\partial (\rho h)}{\partial t} &=& \rho \Hnuc + \rho \Hext\enskip .
\end{eqnarray}
  In particular, to evolve the species, we solve the system:
\begin{eqnarray}
\frac{dX_k}{dt} &=& \omegadot_k(\rho,X_k,T)\enskip, \\
\frac{dT}{dt} &=&\frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right )
\enskip. \label{eq:reacttemp}
\end{eqnarray}
using the stiff ordinary differential equation integration methods
provided by the {\tt VODE} package (cite vode).  The absolute error
tolerances are set to $10^{-12}$ for the species, and a relative
tolerance of $10^{-5}$ is used for the temperature.  The integration
yields the new values of the mass fractions, $X_k^{\outp}$.  Equation
(\ref{eq:reacttemp}) is derived from the full temperature evolution equation
from equation (8) in paper III,
\begin{equation}
\label{eq:tempenthalpy}
\frac{DT}{Dt} = \frac{1}{\rho c_p} \left[ \left(1 - \rho h_p\right) \frac{D p}{D t}
 - \sum_k \rho \xi_k {\omegadot}_k 
 + \rho \Hnuc + \rho \Hext \right] \enskip , 
\end{equation}
by assuming that the pressure is constant during the burn state.  In
evolving these equations, we need to evaluate $c_p$ and $\xi_k$.  In
theory, this means evaluating the equation of state for each
right-hand side evaluation that {\tt VODE} requires.  In practice, we
freeze $c_p$ and $\xi_k$ at the start of the integration time step and
compute them using $\rho^{\inp}, X_k^{\inp},$ and $T^{\inp}$ as inputs
to the equation of state.  We update temperature using
$T^{\outp} = T(\rho^\outp,h^\outp,X_k^\outp)$ if {\tt use\_tfromp = FALSE} or 
$T^{\outp} = T(\rho^\outp,p_0^\inp,X_k^\outp)$  if {\tt use\_tfromp = TRUE}.  
Note that the density remains unchanged
during the {\bf React State} step, i.e., $\rho^{\outp} = \rho^{\inp}$.

@@

\item[planar:] We discretize the base state density evolution 
equation from paper III equation (29) without the $\etarho$ term,
\begin{equation}
\frac{\partial \rhozero}{\partial t} = - \nabla \cdot \left( \rhozero w_0 \eb_r \right)
\enskip ,
\label{eq:rho0upd_new}
\end{equation}
to compute the new base state density,
\begin{equation}
\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp} - \frac{\dt}{\dr} 
\left [ \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j+\myhalf} -  
         \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j-\myhalf} \right ] \enskip,
\end{equation}
where $j$ refers to the one-dimensional index in the radial direction.
The interface states, $\rho_0^{\outp,\nph}$, are computed using the
procedure described in paper II, Appendix A.\\

@@

We discretize the base state enthalpy equation,
\begin{equation}
\frac{\partial (\rho h)_0}{\partial t} = -\nabla\cdot\left[(\rho h)_0w_0\eb_r\right] 
+ \psi \enskip ,
\end{equation}
to compute the new base state enthalpy,
\begin{equation}
(\rho h)_{0,j}^{\outp} = (\rho h)_{0,j}^{\inp} - \frac{\dt}{\Delta r}
\left\{ \left[ (\rho h)_0^{\nph} w_0^{\inp}\right]_{j+\myhalf} -  
         \left[ (\rho h)_0^{\nph} w_0^{\inp}\right]_{j-\myhalf} \right\}
+ \dt\psi_j^{\inp} \enskip ,
\end{equation}
where the interface states $(\rho h)_0^{n+\myhalf}$, are again found via the procedure 
described in paper II, Appendix A.\\

@@

The local velocity field is described by paper III equation (37),
\begin{equation}
\frac{\partial\Ubt}{\partial t} = 
- \left(\Ubt+w_0\right) \cdot \nabla \Ubt
- \left(\Ubt \cdot \eb_r\right) \frac{\partial w_0}{\partial r} \eb_r
- \frac{1}{\rho} \nabla\pi
+ \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} \eb_r
- \frac{(\rho-\rhozero)}{\rho} \; g \; \eb_r  \label{eq:utildeupd}  \enskip .
\end{equation}

We construct time-centered velocities at edges, $\uadvone$, 
using a second-order unsplit Godunov procedure followed by an elliptic
solve, as described in Appendix B of paper III.  We note that 
$\uadvone$ satisfies the discrete versions of paper III equation (21),
\begin{equation}
\int_{\Omega_H} \Ubt \cdot \eb_r \; dA = 0 \label{eq:udoterzero} \enskip ,
\end{equation}
and paper III equation (39),
\begin{equation}
\nabla \cdot (\beta_0 \Ubt )  = \beta_0 \left(S - \Sbar \right )\enskip ,
\label{eq:tildeconstraint}
\end{equation}
with the discretization
\begin{equation}
\nabla \cdot \left(\beta_0^n \uadvone\right) = 
\beta_0^n \left(S^{\nph,\star} - \overline{S^{\nph,\star}}\right) \enskip .
\end{equation}
The discretization with the volume discrepancy correction is
\begin{equation}
\nabla \cdot \left(\beta_0^n \uadvone\right) = 
\beta_0^n \left\{ \left(S^{\nph,\star} - \overline{S^{\nph,\star}}\right)
+ \frac{f}{\gammabar^n p_0^n}
\left[\frac{p(\rho,h,X_k)^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n}\right]\right\} \enskip .
\end{equation}

@@

%--------------------------------------------------------------------------
% STEP 4
%--------------------------------------------------------------------------
\item[Step 4.] {\em Advect the base state and full state, through a time interval of $\dt.$}

@@

\item \label{step:species update} 
Update the species.  Here we consider only the advection terms, 
neglecting the reaction terms, and solve a discretized version of
\begin{equation}
\frac{\partial (\rho X_k)}{\partial t} + \nabla \cdot (\Ub \rho X_k) = 0 \enskip .
\label{eq:species}
\end{equation}
The update consists of two steps:

@@

  \item Compute the species edge states, $(\rho X_k)^{(1),\nph,\star}$,
  for the conservative update of $(\rho X_k)^{(1)}$. 
  Here we predict $\rho^{'(1)} = \rho^{(1)} - \rhozero^n$ and 
  $X_k^{(1)} = (\rho X_k)^{(1)} / \rho^{(1)}$ to time-centered edges to obtain 
  $\rho^{'(1),\nph,\star}$ and $X_k^{(1),\nph,\star}$ using a second-order
  unsplit Godunov procedure, as described in paper II,
  Appendix A, using $\Vb = \uadvone+w_0^{\nph,\star} \eb_r$.  Convert the 
  perturbational edge state to a full state using
\begin{equation}
\rho^{(1),\nph,\star} = 
\rho^{'(1),\nph,\star} + \frac{\rho_0^n + \rho_0^{(2a),\star}}{2} \enskip .
\label{eq:rhoXedgestate}
\end{equation}
  In planar geometry, we use 4th-order spatial averaging to compute $\rho_0$ on 
  radial faces.  In spherical geometry, we first map $\rho_0$ to Cartesian cell-centers 
  using the mapping in section \ref{Sec:1D Cell-Centered to Cartesian Cell-Centered}, 
  and then use 2nd-order spatial averaging to compute $\rho_0$ on all faces.

  \item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2),\star}$
  without explicitly including the reaction terms,
\begin{equation}
(\rho X_k)^{(2),\star} = (\rho X_k)^{(1)} 
 - \dt \; \left\{ \nabla \cdot \left[ \left(\uadvone+w_0^{\nph,\star} \eb_r\right)  
  (\rho X_k)^{(1),\nph,\star} \right] \right\} \enskip ,
\end{equation}
\begin{equation}
\rho^{(2),\star} = \sum_k (\rho X_k)^{(2),\star} \enskip ,
\end{equation}
\begin{equation}
X_k^{(2),\star} = (\rho X_k)^{(2),\star} / \rho^{(2),\star}
\end{equation}

@@

\begin{description}
\item For planar geometry,
\begin{equation}
 \etarho^{\nph,\star} =  {\rm {\bf Avg}} \sum_k \left[ \left(\uadvone \cdot \eb_r + w_0^{\nph,\star}\right) (\rho X_k)^{(1),\nph,\star} \right] - w_0^{\nph,\star} \rho_0^{\nph,\star,{\rm pred}} \enskip ,
\end{equation}
\item For spherical geometry, first construct 
$\eta_{\rho}^{{\rm cart},\nph,\star} = [\rho'(\Ubt\cdot\eb_r)]^{n+\myhalf,\star}$ using:
\begin{equation}
\eta_{\rho}^{{\rm cart},\nph,\star} = \left[\left(\frac{\rho^{(1)}+\rho^{(2),\star}}{2}\right)-\left(\frac{\rho_0^n+\rho_0^{(2a),\star}}{2}\right)\right] \sum_d \left(\uadvone \cdot \eb_d\right)n_d.
\end{equation}
Then, $\etarho^{\nph,\star}$ is the cell-centered average of $\eta_{\rho}^{{\rm cart},\nph,\star}$,
\begin{equation}
\etarho^{\nph,\star} = \overline{\eta_{\rho}^{{\rm cart},\nph,\star}}.
\end{equation}
This gives a cell-centered $\etarho^{\nph,\star}$.  To get $\etarho^{\nph,\star}$ at 
radial edges, average the two neighboring radial cell-centers.
\end{description}

@@

\item Update the enthalpy.  Here we consider only the advective terms, 
neglecting the reaction terms and thermal diffusion terms, and solve a discretized version of
  \begin{equation}
  \frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\Ub \rho h)
+ \psi + (\Ubt \cdot \eb_r) \frac{\partial p_0}{\partial r} 
\enskip , \label{eq:rhohupdadv} 
  \end{equation}
For spherical geometry, experience has shown that solving the analytically equivalent form,
\begin{equation}
  \frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\Ub \rho h)
+ \psi + \nabla \cdot (\Ubt p_0) - p_0 \nabla \cdot \Ubt  
\enskip , \label{eq:rhohupdadv2} 
\end{equation}
results in a better solution.

@@

  \item Compute the enthalpy edge state, $(\rho h)^{(1),\nph,\star},$
    for the conservative update of $(\rho h)^{(1)}.$  Here we predict 
    $(\rho h)' = (\rho h)^{(1)} - (\rho h)_0$ to time-centered edges to obtain 
    $(\rho h)^{'(1),\nph,\star}$, 
    using a second-order unsplit Godunov procedure, 
    as described in paper II, Appendix A, using $\Vb =
    \uadvone+w_0^{\nph,\star} \eb_r$.  We do not include the reaction
    terms in the enthalpy prediction, since we accounted for them
    already in {\bf React State}.  Convert the perturbational edge
    state to a full state using

@@

  \begin{eqnarray}
  (\rho h)^{(2),\star} &=& (\rho h)^{(1)} - \dt \; \left\{ \nabla
      \cdot \left[ \left(\uadvone+w_0^{\nph,\star} \eb_r\right) (\rho
      h)^{(1),\nph,\star} \right] \right\} \nonumber \\ && + \dt \;
    \left(\uadvone \cdot \eb_r\right) \left(\frac{\partial
      p_0}{\partial r} \right)^{n} + \dt \; \psi^{\nph,\star} \enskip ,
  \end{eqnarray}

@@

  \begin{eqnarray}
  (\rho h)^{(2),\star} &=& (\rho h)^{(1)} - \dt \; \left\{ \nabla
      \cdot \left[ \left(\uadvone+w_0^{\nph,\star} \eb_r\right) (\rho
      h)^{(1),\nph,\star} \right] \right\} \nonumber \\ 
    && + \dt \; \left \{ \nabla \cdot \left (\uadvone p_0^{n} \right ) 
       - p_0^{n} \nabla \cdot \uadvone \right \} \nonumber \\
    && + \dt \; \psi^{\nph,\star} \enskip ,
  \end{eqnarray}

@@

The procedure to construct $\uadvtwo$ is described in detail in Appendix B of paper III
and is analogous to the procedure used in step 2, but with updated values
for $w_0$ and $\pizero.$  We note that $\uadvtwo$ satisfies the discrete versions of 
equations (\ref{eq:udoterzero}) and (\ref{eq:tildeconstraint}), with the discretization
\begin{equation}
\nabla \cdot \left(\beta_0^{\nph} \uadvtwo\right) =
\beta_0^{\nph}\left(S^{\nph} - \overline{S^{\nph}}\right) \enskip ,
\end{equation}
\begin{equation}
\beta_0^{\nph} = \frac{ \beta_0^n +  \beta_0^{n+1,\star} }{2}; ~~~~~
 \beta_0^{n+1,\star} = \beta \left(\rho_0^{n+1,\star}, p_0^{n+1,\star}, \gammabar^{n+1,\star}, g^{n+1,\star}\right) \enskip ,
\end{equation}
where $\beta_0$ is computed as described in paper III, Appendix C.

The discretization with the volume discrepancy correction is
\begin{equation}
\nabla \cdot \left(\beta_0^{\nph} \uadvtwo\right) =
\beta_0^{\nph}\left\{\left(S^{\nph} - \overline{S^{\nph}}\right)
 + \frac{f}{\gammabar^{\nph} p_0^{\nph}}\cdot 
\left[\frac{p(\rho,h,X_k)^{\nph} - \overline{p(\rho,h,X_k)^{\nph}}}{\Delta t^n}\right]\right\} \enskip ,
\end{equation}

%--------------------------------------------------------------------------
% STEP 8
%--------------------------------------------------------------------------
\item[Step 8.] {\em Advect the base state and full state through a time interval of $\dt.$}

@@

\item Compute the species edge states, $(\rho X_k)^{(1),n+\myhalf}$, 
using $\Vb = \uadvtwo+w_0^{\nph}\eb_r$.  Convert the perturbational edge 
state to a full state using 
\begin{equation}
\rho^{(1),\nph} = \rho^{'(1),\nph} + \frac{\rho_0^n + \rho_0^{(2a)}}{2}\enskip .
\end{equation}

\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2)}$ using
\begin{equation}
(\rho X_k)^{(2)} = (\rho X_k)^{(1)} 
- \dt \; \left\{ \nabla \cdot \left[\left(\uadvtwo+w_0^{\nph} \eb_r\right)  
(\rho X_k)^{(1),\nph} \right] \right\}  \enskip ,
\end{equation}
\begin{equation}
\rho^{(2)} = \sum_k (\rho X_k)^{(2)} \enskip ,
\end{equation}
\begin{equation}
X_k^{(2)} = (\rho X_k)^{(2)} / \rho^{(2)}
\end{equation}

@@

\item For planar geometry,
\begin{equation}
 \etarho^{\nph} = {\rm {\bf Avg}} \sum_k \left [\left(\uadvtwo \cdot \eb_r + w_0^{\nph}\right) (\rho X_k)^{(1),\nph} \right] - w_0^{\nph} \rho_0^{\nph,{\rm pred}} \enskip ,
\end{equation}
\item For spherical geometry, first construct 
$\eta_{\rho}^{{\rm cart},\nph} = [\rho'(\Ubt\cdot\eb_r)]^{n+\myhalf}$ using:
\begin{equation}
\eta_{\rho}^{{\rm cart},\nph} = \left[\left(\frac{\rho^{(1)}+\rho^{(2)}}{2}\right)-\left(\frac{\rho_0^n+\rho_0^{(2a)}}{2}\right)\right] \sum_d \left(\uadvtwo \cdot \eb_d\right)n_d.
\end{equation}
Then, $\etarho^{\nph}$ is the cell-centered average of $\eta_{\rho}^{{\rm cart},\nph}$,
\begin{equation}
\etarho^{\nph} = \overline{\eta_{\rho}^{{\rm cart},\nph}}.
\end{equation}
This gives a cell-centered $\etarho^{\nph}$.  To get $\etarho^{\nph}$ at 
radial edges, average the two neighboring cell-centers.
\end{description}
